Notes on Matrix Factorization for Hartree-Fock Stability of HEG

### 2017-04-05 doc/papers/matrixproperties/MatrixProperties.tex
Checkpoint for similarity transform explanation
Added cRHF2cRHF (singlet instability) functionality

\documentclass{revtex4}
\usepackage[margin=1in, paperheight=1in, paperwidth=1in]{geometry}  % doesn't compile w/o width, height
\geometry{
 a4paper,
 total={170mm,257mm},
 left=15mm,
 top=5mm,
}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{kbordermatrix}
\usepackage{braket}
\usepackage{xcolor}
\usepackage{graphicx}

@@

\begin{document}
\title{Notes on Matrix Factorization for Hartree-Fock Stability of HEG}
\author{Evan Curtin}
\maketitle

\section{Orbital Hessian Factorization}
According to Seeger and Pople\cite{Seeger1977}, (and many other sources) the molecular orbital
Hessian has the form,
\begin{eqnarray}\label{basic}
\mathbf{H} =
  \begin{bmatrix}
    \mathbf{A}   & \mathbf{B}   \\
    \mathbf{B^*} & \mathbf{A^*} \\
  \end{bmatrix}
\end{eqnarray}
Where the matrices denoted by $\mathbf{A}$ and $\mathbf{B}$ are given by,
\begin{eqnarray}
  A_{st} &=& \e\diag +  \textcolor{myred}{\braket{aj||ib}} \\
  B_{st} &=& \textcolor{myblue}{\braket{ab||ij}}         
\end{eqnarray}
and the two electron integral used here is defined as, 
\begin{eqnarray}
  \braket{pq|rs} &=& \int\int \chi^*_p(1) \chi^*_q(2) \frac{1}{r_{12}} \chi^*_r(1) 
  \chi^*_s(2)
  d\tau d\sigma \\
  \braket{pq||rs} &=& \braket{pq|rs} - \braket{pq|sr} \\
  (pq|rs) &=& \int\int \chi^*_p(1) \chi^*_q(2) \frac{1}{r_{12}} \chi^*_r(1) 
  \chi^*_s(2)
  d\tau \\
  (pq||rs) &=& (pq|rs) - (pq|sr)
\end{eqnarray}
Where $\sigma$ is the spin coordinate and $d\tau$ is the volume element in all spatial coordinates.
Additionally, the convention of $a, b$ being virtual states, $i, j$ being occupied states and 
$p, q, r, s$ being any state is followed.

The solution is said to be unstable when the lowest eigenvalue is strictly negative. The
corresponding eigenvector points downhill in energy in the case of instability. Defining
the eigenvector as $\mathbf{D}$, the eigenvalue problem can be written as
\begin{eqnarray}
  \mathbf{HD = \lambda D}
\end{eqnarray}
In many cases, the matrix $\mathbf{H}$ factorizes into various components. The following
form of a similarity transformed eigenvalue problem is useful to keep in mind:
\begin{eqnarray}
  \mathbf{S^{-1} H S S^{-1} D} &=& \mathbf{\lambda S^{-1} D} \\
  \mathbf{\tilde H \tilde D} &=& \mathbf{\lambda \tilde D}   
\end{eqnarray}

Thus if the matrix $\mathbf{H}$ can be transformed via a similarity transformation defined
by $\mathbf{S}$ into $\mathbf{\tilde H}$, then the corresponding eigenvector which satisfies the 
eigenvalue problem
has the form $\mathbf{\tilde D} = \mathbf{S^{-1}D}$.  
Writing the eigenvalue problem in the form of Seeger \& Pople (eq. 18 in \cite{Seeger1977}), 
we get
\begin{eqnarray}
  \begin{bmatrix}
    \mathbf{A} & \mathbf{B} \\
    \mathbf{B^*} & \mathbf{A^*} 
  \end{bmatrix}
  \begin{bmatrix}
    \mathbf{d} \\
    \mathbf{d^*} 
  \end{bmatrix}
  = 2E_2
  \begin{bmatrix}
    \mathbf{d} \\
    \mathbf{d^*} 
  \end{bmatrix}
\end{eqnarray}
We can now apply the similarity transform defined by the Unitary matrix 
\begin{eqnarray}
  \mathbf{U} = \frac{1}{\sqrt{2}}
  \begin{bmatrix}
    \mathbf{I} & -\mathbf{I} \\
    \mathbf{I} & \mathbf{I} 
  \end{bmatrix}
\end{eqnarray}
after which the transformed eigenvalue problem has the form
\begin{eqnarray}
  \frac{1}{2}
  \begin{bmatrix}
    \mathbf{A + B + A^* + B^*} & \mathbf{-A + A^* + B - B^*} \\
    \mathbf{-A + A^* - B + B^*} & \mathbf{A^* + A - B - B^*} 
  \end{bmatrix}
  \begin{bmatrix}
    \mathbf{d + d^*} \\
    \mathbf{d - d^*} 
  \end{bmatrix}
  &=& 2E_2
  \begin{bmatrix}
    \mathbf{d + d^*} \\
    \mathbf{d - d^*} 
  \end{bmatrix} \\
  &=& 2E_2
  \begin{bmatrix}
    \mathbf{Re(d)} \\
    \mathbf{Im(d)} 
  \end{bmatrix}
\end{eqnarray}
If $\mathbf{A}$ and $\mathbf{B}$ are both real, $\mathbf{A = A^*}$ and $\mathbf{B = B^*}$ and the 
above simplifies to 
\begin{eqnarray}
  \begin{bmatrix}
    \mathbf{A + B} & \mathbf{0} \\
    \mathbf{0} & \mathbf{A - B} 
  \end{bmatrix}
  \begin{bmatrix}
    \mathbf{Re(d)} \\
    \mathbf{Im(d)} 
  \end{bmatrix}
  &=& 2E_2   
  \begin{bmatrix}
    \mathbf{Re(d)} \\
    \mathbf{Im(d)} 
  \end{bmatrix}
\end{eqnarray}
Clearly, the matrix factorizes into $\mathbf{A+B}$ and $\mathbf{A-B}$. The eigenvectors factorize
into the real and imaginary displacements of the orbitals, respectively. This is equivalent to the 
condition derived in and around eq. 20 for internal and external instabilities of real
GHF solutions \cite{Seeger1977}, however this approach applies to any matrix with the same
form as equation \ref{basic}.

\section{Forms of equations with UHF wavefunction}
In the case of a stationary UHF solution, the matrices $\mathbf{A}$ and $\mathbf{B}$ have the
following forms, after integrating over spin:
\begin{eqnarray}
  A_{st} &=& \e\diag \textcolor{myred}{+ \delta_{\sigma_i\sigma_a}\delta_{\sigma_j\sigma_b}\Ap
                     - \delta_{\sigma_a\sigma_b}\delta_{\sigma_i\sigma_j}\App} \\
  B_{st} &=& \textcolor{myblue}{\delta_{\sigma_a\sigma_i}\delta_{\sigma_b\sigma_j}\B 
             - \delta_{\sigma_a\sigma_j}\delta_{\sigma_b\sigma_i}\Br} 
\end{eqnarray}
In matrix form, 
\begin{eqnarray}
  \mathbf{A}&=&\kbordermatrix{
        & \AtoA           & \AtoB           & \BtoA          & \BtoB          \\
  \AtoA & \e\diag + \Aa   & 0               & 0              & \Ap            \\
  \AtoB & 0               & \e\diag - \App  & 0              & 0              \\
  \BtoA & 0               & 0               & \e\diag - \App & 0              \\
  \BtoB & \Ap             & 0               & 0              & \e\diag + \Aa  
}
\end{eqnarray}
\begin{eqnarray}
  \mathbf{B}&=&\kbordermatrix{
        & \AtoA           & \AtoB           & \BtoA          & \BtoB          \\
  \AtoA & \Ba             & 0               & 0              & \B             \\
  \AtoB & 0               & 0               & -\Br           & 0              \\
  \BtoA & 0               & -\Br            & 0              & 0              \\
  \BtoB & \B              & 0               & 0              & \Ba            
}
\end{eqnarray}
\\
These matrices factorize into ``spin conserved'' $(\mathbf{A', B'})$ and
``spin-unconserved'' $(\mathbf{A'', B''})$ parts, to use the language of Seeger
and Pople. The spin conserved matrices are given by

@@

\begin{eqnarray}
  \mathbf{B''}&=&\kbordermatrix{
        & \AtoB           & \BtoA  \\
  \AtoB & 0               & -\Br   \\
  \BtoA & -\Br            & 0      \\
}
\end{eqnarray}
\\
Thus the spin conserved molecular orbital hessian, 
\begin{eqnarray}
\mathbf{H}' =
  \begin{bmatrix}
    \mathbf{A}'   & \mathbf{B}'   \\
    \mathbf{B}'^* & \mathbf{A}'^* \\
  \end{bmatrix}
\end{eqnarray}
is given by
\begin{eqnarray*}
  \mathbf{H'}&=&\kbordermatrix{
        & \AtoA             & \BtoB            & \AtoA             & \BtoB            \\
  \AtoA & \e\diag + \Aa     & \Ap              & \Ba               & \B               \\
  \BtoB & \Ap               & \e\diag + \Aa    & \B                & \Ba              \\
  \AtoA & \Ba^*             & \B^*             & \e\diag + \Aa^*   & \Ap^*            \\
  \BtoB & \B^*              & \Ba^*            & \Ap^*             & \e\diag + \Aa^*  \\
}
\end{eqnarray*}
\\
and the spin unconserved molecular orbital hessian, 
\begin{eqnarray}
\mathbf{H}'' =
  \begin{bmatrix}
    \mathbf{A}''   & \mathbf{B}''   \\
    \mathbf{B}''^* & \mathbf{A}''^* \\
  \end{bmatrix}
\end{eqnarray}
is given by:
\begin{eqnarray*}
  \mathbf{H''}&=&\kbordermatrix{
        & \AtoB           & \BtoA              & \AtoB             & \BtoA            \\
  \AtoB & \e\diag - \App  & 0                  & 0                 & -\Br             \\
  \BtoA & 0               & \e\diag - \App     & -\Br              & 0                \\
  \AtoB & 0                 & -\Br^*           & \e\diag - \App^*  & 0                \\
  \BtoA & -\Br^*            & 0                & 0                 & \e\diag - \App^* \\
}
\end{eqnarray*}
\\
The eigenvalue equation defining $\mathbf{D}''$ is, 
\begin{eqnarray}
  \mathbf{H''D''} = 
  \mathbf{H''}
  \begin{bmatrix}
    \mathbf{d_\AtoB} \\
    \mathbf{d_\BtoA} \\
    \mathbf{d^*_\AtoB} \\        
    \mathbf{d^*_\BtoA} 
  \end{bmatrix}
  = 2E_2
  \begin{bmatrix}
    \mathbf{d_\AtoB} \\
    \mathbf{d_\BtoA} \\
    \mathbf{d^*_\AtoB} \\        
    \mathbf{d^*_\BtoA} 
  \end{bmatrix}
\end{eqnarray}

Transforming the equation with the similarity transform defined by the following Unitary 
matrix, 
\begin{eqnarray}
  \mathbf{U} = \frac{1}{\sqrt{2}}
  \begin{bmatrix}
    \mathbf{I} & 0 & 0 & -\mathbf{I} \\
     0 & \mathbf{I} & -\mathbf{I} & 0 \\ 
     0 & \mathbf{I} & \mathbf{I} & 0 \\
    \mathbf{I} & 0 & 0 & \mathbf{I} \\  
  \end{bmatrix}
\end{eqnarray}
The transformed Hessian is ($e = \e\diag$, $C.C.$ is the complex conjugate of the rest of the 
expression). 
\begin{eqnarray*}
\hspace{-1.45cm}
  &\mathbf{\tilde{H}''}& = \\
  \hspace{-1.45cm} 
  &\frac{1}{2}& \kbordermatrix{
        & \AtoB           & \BtoA              & \AtoB             & \BtoA            \\
  \AtoB & e - \App -\Br + C.C.  & 0 & 0  & -e + \App -\Br - C.C.             \\
  \BtoA & 0               & e - \App -\Br + C.C.     & -e + \App -\Br - C.C. & 0  \\
  \AtoB & 0 & -e + \App +\Br - C.C. & e - \App +\Br + C.C.  & 0                \\
  \BtoA & -e + \App +\Br - C.C.    & 0  & 0  & e - \App +\Br +  C.C. \\
}
\end{eqnarray*}
While the corresponding transformed eigenvector has the following form:
\begin{eqnarray*}
  \mathbf{\tilde{D}''} = 
  \begin{bmatrix}
    d_\AtoB + d^*_\BtoA \\
    d_\BtoA + d^*_\AtoB \\
    -d_\BtoA + d^*_\AtoB \\
    -d_\AtoB + d^*_\BtoA \\
  \end{bmatrix}
\end{eqnarray*}

If the two electron integrals and energies are all real, we have for the form 
of $\mathbf{H''}$, 
\begin{eqnarray*}
\hspace{-1cm}
  \mathbf{\tilde{H}''} = 
  \kbordermatrix{
        & \AtoB           & \BtoA              & \AtoB             & \BtoA        \\
  \AtoB & \e - \App -\Br  & 0                  & 0                 & 0            \\
  \BtoA & 0               & \e - \App -\Br     & 0                  & 0  \\
  \AtoB & 0 & 0 & \e - \App + \Br  & 0                \\
  \BtoA & 0    & 0  & 0  &\e - \App + \Br \\
}
\end{eqnarray*}
Furthermore, if the solution we started with was an RHF solution, the $\AtoB$ and 
$\BtoA$ are identical. Letting $d_\AtoB = d_\BtoA = d$, we get the following forms 
of the transformed eigenvectors, 
\begin{eqnarray}\label{rhfvecs}
  \mathbf{\tilde{D}''} = 
  \begin{bmatrix}
    d + d^*  \\
    d + d^*  \\
    -d + d^* \\
    -d + d^* \\
  \end{bmatrix}
  =
  \begin{bmatrix}
    Re(d)  \\
    Re(d)  \\
    Im(d)  \\
    Im(d)  \\
  \end{bmatrix}
\end{eqnarray}
At this point the only assumptions we made are that the two electron integrals and the
energies are real-valued and that the solution we started with was an RHF solution. From
this it immediately follows that the eigenvectors have the form of equation \ref{rhfvecs}.

Clearly, the eigensystem defined by $\mathbf{H}''\mathbf{D}'' = 2E_2\mathbf{D}''$ is a
doubly degenerate set of the following equations:
\begin{eqnarray}
  \left[ \mathbf{A+B} \right] Re(d) &=& 2E_2 Re(d) \\
  \left[ \mathbf{A-B} \right] Im(d) &=& 2E_2 Im(d) 
\end{eqnarray}
Where $\mathbf{A}$ and $\mathbf{B}$ are defined such that
\begin{eqnarray}
  A_{ia, jb} &=& \e\diag - \App \\
  B_{ia, jb} &=& -\Br 
\end{eqnarray}
These matrices are exactly equivalent to ${}^3\mathbf{A}'$ and ${}^3\mathbf{B}'$ (eq 35 in Seeger
\& Pople). Thus, when the matrices $\mathbf{A}$ and $\mathbf{B}$ are real, and the solution
is an RHF solution, the eigenvalue problem is doubly degenerate, and factorizes into the equation
characterized by the triplet instability equation.

\section{Proof of Real-Valued A and B}
In the case of the Homogeneous electron gas,
the two electron integral is given by
(eq. 12 of~\cite{Delyon2008} and p. 16 of~\cite{Guiliani2005}):

@@

So far I have not used this to simplify anything, but it is true.

\section{References}
\bibliography{../heg_references}
\end{document}

>>>>

\documentclass{revtex4}
\usepackage[margin=1in, paperheight=1in, paperwidth=1in]{geometry}  % doesn't compile w/o width, height
\geometry{
 a4paper,
 total={170mm,257mm},
 left=15mm,
 top=5mm,
}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{kbordermatrix}
\usepackage{braket}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{verbatim}

@@

\begin{document}
\title{Notes on Matrix Factorization for Hartree-Fock Stability of HEG}
%\author{Evan Curtin}
\maketitle

\section{Definitions}
According to Seeger and Pople\cite{Seeger1977}, (and many other sources) the molecular orbital
Hessian has the form,
\begin{eqnarray}\label{basic}
\mathbf{H} =
  \begin{bmatrix}
    \mathbf{A}   & \mathbf{B}   \\
    \mathbf{B^*} & \mathbf{A^*} \\
  \end{bmatrix}
\end{eqnarray}
Where the matrices denoted by $\mathbf{A}$ and $\mathbf{B}$ are given by,
\begin{eqnarray}
  A_{st} &=& \e\diag +  \textcolor{myred}{\braket{aj||ib}} \\
  B_{st} &=& \textcolor{myblue}{\braket{ab||ij}}         
\end{eqnarray}
and the two electron integral used here is defined as, 
\begin{eqnarray}
  \braket{pq|rs} &=& \int\int \chi^*_p(1) \chi^*_q(2) \frac{1}{r_{12}} \chi_r(1) 
  \chi_s(2)
  d\tau d\sigma \\
  \braket{pq||rs} &=& \braket{pq|rs} - \braket{pq|sr} \\
  (pq|rs) &=& \int\int \chi^*_p(1) \chi^*_q(2) \frac{1}{r_{12}} \chi_r(1) 
  \chi_s(2)
  d\tau \\
  (pq||rs) &=& (pq|rs) - (pq|sr)
\end{eqnarray}
Where $\sigma$ is the spin coordinate and $d\tau$ is the volume element in all spatial coordinates.
Additionally, the convention of $a, b$ being virtual states, $i, j$ being occupied states and 
$p, q, r, s$ being any state is followed.

The solution is said to be unstable when the lowest eigenvalue of $\mathbf{H}$
is strictly negative. The
corresponding eigenvector points downhill in energy in the case of instability. Defining
the eigenvector as $\mathbf{D}$, the eigenvalue problem can be written as
\begin{eqnarray}
  \mathbf{HD = \lambda D}
\end{eqnarray}

\section{Transforming H Dictates the form of D}
In many cases, the matrix $\mathbf{H}$ factorizes into various components. The following
form of a similarity transformed eigenvalue problem is useful to keep in mind:
\begin{eqnarray}
  \mathbf{S^{-1} H S S^{-1} D} &=& \mathbf{\lambda S^{-1} D} \\
  \mathbf{\tilde H \tilde D} &=& \mathbf{\lambda \tilde D}   
\end{eqnarray}

Thus if the matrix $\mathbf{H}$ can be transformed via a similarity transformation defined
by $\mathbf{S}$ into $\mathbf{\tilde H}$, then the corresponding eigenvector which satisfies the 
eigenvalue problem
has the form $\mathbf{\tilde D} = \mathbf{S^{-1}D}$.

This approach is commonly used in the paper to transform into the basis that separates
the real and complex contributions of the eigenvectors when $\mathbf{A}$ and $\mathbf{B}$ are real.
Writing the eigenvalue problem in the form of Seeger \& Pople (eq. 18 in \cite{Seeger1977}), 
we get
\begin{eqnarray}
  \begin{bmatrix}
    \mathbf{A} & \mathbf{B} \\
    \mathbf{B^*} & \mathbf{A^*} 
  \end{bmatrix}
  \begin{bmatrix}
    \mathbf{d} \\
    \mathbf{d^*} 
  \end{bmatrix}
  = 2E_2
  \begin{bmatrix}
    \mathbf{d} \\
    \mathbf{d^*} 
  \end{bmatrix}
\end{eqnarray}
We can now apply the similarity transform defined by the Unitary matrix 
\begin{eqnarray}
  \mathbf{U} = \frac{1}{\sqrt{2}}
  \begin{bmatrix}
    \mathbf{I} & -\mathbf{I} \\
    \mathbf{I} & \mathbf{I} 
  \end{bmatrix}
\end{eqnarray}
after which the transformed eigenvalue problem has the form
\begin{eqnarray}
  \frac{1}{2}
  \begin{bmatrix}
    \mathbf{A + B + A^* + B^*} & \mathbf{-A + A^* + B - B^*} \\
    \mathbf{-A + A^* - B + B^*} & \mathbf{A^* + A - B - B^*} 
  \end{bmatrix}
  \begin{bmatrix}
    \mathbf{d + d^*} \\
    \mathbf{d - d^*} 
  \end{bmatrix}
  &=& 2E_2
  \begin{bmatrix}
    \mathbf{d + d^*} \\
    \mathbf{-d + d^*} 
  \end{bmatrix} \\
  &=& 2E_2
  \begin{bmatrix}
    \mathbf{Re(d)} \\
    \mathbf{Im(d)} 
  \end{bmatrix}
\end{eqnarray}
If $\mathbf{A}$ and $\mathbf{B}$ are both real, $\mathbf{A = A^*}$ and $\mathbf{B = B^*}$ and the 
above simplifies to 
\begin{eqnarray}
  \begin{bmatrix}
    \mathbf{A + B} & \mathbf{0} \\
    \mathbf{0} & \mathbf{A - B} 
  \end{bmatrix}
  \begin{bmatrix}
    \mathbf{Re(d)} \\
    \mathbf{Im(d)} 
  \end{bmatrix}
  &=& 2E_2   
  \begin{bmatrix}
    \mathbf{Re(d)} \\
    \mathbf{Im(d)} 
  \end{bmatrix}
\end{eqnarray}
Clearly, the matrix factorizes into $\mathbf{A+B}$ and $\mathbf{A-B}$. The eigenvectors factorize
into the real and imaginary displacements of the orbitals, respectively. This is equivalent to the 
condition derived in and around eq. 20 in Seeger/Pople for internal and external instabilities of 
real
GHF solutions \cite{Seeger1977}, however this approach applies to any matrix with the same
form as equation \ref{basic}.

\section{Forms of equations with UHF Solution}
In the case of a stationary UHF solution, the elements of matrices $\mathbf{A}$ and $\mathbf{B}$ 
have the following forms, after integrating over spin:
\begin{eqnarray}
  A_{i \rightarrow a, j \rightarrow b} &=& \e\diag \textcolor{myred}{+ 
  \delta_{\sigma_i\sigma_a}\delta_{\sigma_j\sigma_b}\Ap
                     - \delta_{\sigma_a\sigma_b}\delta_{\sigma_i\sigma_j}\App} \\
  B_{i \rightarrow a,j \rightarrow b} &=& 
  \textcolor{myblue}{\delta_{\sigma_a\sigma_i}\delta_{\sigma_b\sigma_j}\B 
             - \delta_{\sigma_a\sigma_j}\delta_{\sigma_b\sigma_i}\Br} 
\end{eqnarray}
Writing down the matrices, where the row index corresponds to $i \rightarrow a$ excitations  and 
the column index corresponds to $j \rightarrow b$ excitations:
\begin{eqnarray*}
  \mathbf{A}&=&\kbordermatrix{
        & \AtoA           & \AtoB           & \BtoA          & \BtoB          \\
  \AtoA & \e\diag + \Aa   & 0               & 0              & \Ap            \\
  \AtoB & 0               & \e\diag - \App  & 0              & 0              \\
  \BtoA & 0               & 0               & \e\diag - \App & 0              \\
  \BtoB & \Ap             & 0               & 0              & \e\diag + \Aa  
}
\end{eqnarray*}
\begin{eqnarray*}
  \mathbf{B}&=&\kbordermatrix{
        & \AtoA           & \AtoB           & \BtoA          & \BtoB          \\
  \AtoA & \Ba             & 0               & 0              & \B             \\
  \AtoB & 0               & 0               & -\Br           & 0              \\
  \BtoA & 0               & -\Br            & 0              & 0              \\
  \BtoB & \B              & 0               & 0              & \Ba            
}
\end{eqnarray*}
\\
These matrices factorize into ``spin conserved'' $(\mathbf{A', B'})$ and
``spin-unconserved'' $(\mathbf{A'', B''})$ parts, to use the language of Seeger
and Pople. Similarly, $\mathbf{H}$ factorizes, with no approximations, into
$\mathbf{H}'$ and $ \mathbf{H}''$
\begin{eqnarray}
\mathbf{H}' =
  \begin{bmatrix}
    \mathbf{A}'   & \mathbf{B}'   \\
    \mathbf{B}'^* & \mathbf{A}'^* \\
  \end{bmatrix}
\end{eqnarray}
\begin{eqnarray}
\mathbf{H}'' =
  \begin{bmatrix}
    \mathbf{A}''   & \mathbf{B}''   \\
    \mathbf{B}''^* & \mathbf{A}''^* \\
  \end{bmatrix}
\end{eqnarray}
The spin conserved matrices are given by

@@

\begin{eqnarray}
  \mathbf{B''}&=&\kbordermatrix{
        & \AtoB           & \BtoA  \\
  \AtoB & 0               & -\Br   \\
  \BtoA & -\Br            & 0      \\
}
\end{eqnarray}
\\
Thus the spin conserved molecular orbital hessian, is given by
\begin{eqnarray*}
  \mathbf{H'}&=&\kbordermatrix{
        & \AtoA             & \BtoB            & \AtoA             & \BtoB            \\
  \AtoA & \e\diag + \Aa     & \Ap              & \Ba               & \B               \\
  \BtoB & \Ap               & \e\diag + \Aa    & \B                & \Ba              \\
  \AtoA & \Ba^*             & \B^*             & \e\diag + \Aa^*   & \Ap^*            \\
  \BtoB & \B^*              & \Ba^*            & \Ap^*             & \e\diag + \Aa^*  \\
}
\end{eqnarray*}
\\
and the spin unconserved molecular orbital hessian, is given by:
\begin{eqnarray*}
  \mathbf{H''}&=&\kbordermatrix{
        & \AtoB           & \BtoA              & \AtoB             & \BtoA            \\
  \AtoB & \e\diag - \App  & 0                  & 0                 & -\Br             \\
  \BtoA & 0               & \e\diag - \App     & -\Br              & 0                \\
  \AtoB & 0                 & -\Br^*           & \e\diag - \App^*  & 0                \\
  \BtoA & -\Br^*            & 0                & 0                 & \e\diag - \App^* \\
}
\end{eqnarray*}

The eigenvalue equation defining $\mathbf{D}'$ is, 
\begin{eqnarray}\label{hppeval}
  \mathbf{H'D'} = 
  \mathbf{H'}
  \begin{bmatrix}
    \mathbf{d_\AtoA} \\
    \mathbf{d_\BtoB} \\
    \mathbf{d^*_\AtoA} \\        
    \mathbf{d^*_\BtoB} 
  \end{bmatrix}
  = 2E_2
  \begin{bmatrix}
    \mathbf{d_\AtoA} \\
    \mathbf{d_\BtoB} \\
    \mathbf{d^*_\AtoA} \\        
    \mathbf{d^*_\BtoB} 
  \end{bmatrix}
\end{eqnarray}

Applying the transformation defined by the unitary matrix, 
\begin{eqnarray}
  \mathbf{U} = \frac{1}{\sqrt{2}} 
  \begin{bmatrix}
    \mathbf{1} & -\mathbf{1} & 0 &  0 \\
    \mathbf{1} &  \mathbf{1} & 0 &  0 \\
    0 &  0 & \mathbf{1} & -\mathbf{1} \\
    0 &  0 & \mathbf{1} &  \mathbf{1} \\
  \end{bmatrix}
\end{eqnarray}

With no further assumptions, this transformation doesn't do much. However, if the spatial
orbitals corresponding to $\alpha$ and $\beta$ are assumed equivalent, the 
transformed $\mathbf{H}'$ is
\begin{eqnarray*}
  \hspace{-1cm}
  \mathbf{U^{-1}H'U} =
  \begin{bmatrix}
    \e + \Aa + \Ap & 0              & \Ba + \B            & 0            \\
    0                    & \e + \Aa - \Ap    & 0                & \Ba - \B              \\
    \Ba^* + \B^*         & 0                & \e+ \Aa^* + \Ap^*   & 0                     \\
    0                    & \Ba^* - \B^*     & 0                 & \e+ \Aa^* - \Ap^*  \\
  \end{bmatrix}
\end{eqnarray*}

while the corresponding eigenvectors are 
\begin{eqnarray*}
  \mathbf{U^{-1}D'} = \frac{1}{\sqrt{2}}
  \begin{bmatrix}
    \mathbf{d_\AtoA + d_\BtoB} \\
    \mathbf{d_\AtoA - d_\BtoB} \\
    \mathbf{d^*_\AtoA + d^*_\BtoB} \\        
    \mathbf{d^*_\AtoA - d^*_\BtoB} 
  \end{bmatrix}
\end{eqnarray*}

In this basis the equations clearly factorize into the following sets of eigenvalue equations:  
\begin{eqnarray}\label{1Heval}
  \begin{bmatrix}
    \e + \Aa + \Ap & \Ba + \B          \\
    \Ba^* + \B^*   & \e+ \Aa^* + \Ap^* \\
  \end{bmatrix}
  \begin{bmatrix}
    \mathbf{d_\AtoA + d_\BtoB} \\
    \mathbf{d^*_\AtoA + d^*_\BtoB} \\        
  \end{bmatrix}
  = 2E_2
  \begin{bmatrix}
    \mathbf{d_\AtoA + d_\BtoB} \\
    \mathbf{d^*_\AtoA + d^*_\BtoB} \\        
  \end{bmatrix}
  \vspace{5mm}
  \\ \label{3Heval}
  \begin{bmatrix}
    \e + \Aa - \Ap & \Ba - \B          \\
    \Ba^* - \B^*   & \e+ \Aa^* - \Ap^* \\
  \end{bmatrix}
  \begin{bmatrix}
    \mathbf{d_\AtoA - d_\BtoB} \\
    \mathbf{d^*_\AtoA - d^*_\BtoB} \\        
  \end{bmatrix}
  = 2E_2
  \begin{bmatrix}
    \mathbf{d_\AtoA - d_\BtoB} \\
    \mathbf{d^*_\AtoA - d^*_\BtoB} \\        
  \end{bmatrix}
\end{eqnarray}
Equation \ref{3Heval} defines the matrix ${{}^3\mathbf{H}'}$, the triplet instability matrix,
while \ref{1Heval} defines the matrix ${{}^1\mathbf{H}'}$, which defines instabilities that mix 
singlet states. The only assumption made until this point in the analysis of ${\mathbf{H}'}$ was
that the original solution was a complex RHF solution. No assumption was made about the outside 
space towards which the eigenvector points. Even still, the eigenvectors can be shown to have the
form given in equations \ref{1Heval} and \ref{3Heval}. Thus, the space in which the eigenvectors
of ${\mathbf{H}'}$ point can only be that of UHF solutions.

Now, the other part of the general form of the stability matrix is ${\mathbf{H}''}$.
The eigenvalue equation defining $\mathbf{D}''$ is, 
\begin{eqnarray}\label{hppeval}
  \mathbf{H''D''} = 
  \mathbf{H''}
  \begin{bmatrix}
    \mathbf{d_\AtoB} \\
    \mathbf{d_\BtoA} \\
    \mathbf{d^*_\AtoB} \\        
    \mathbf{d^*_\BtoA} 
  \end{bmatrix}
  = 2E_2
  \begin{bmatrix}
    \mathbf{d_\AtoB} \\
    \mathbf{d_\BtoA} \\
    \mathbf{d^*_\AtoB} \\        
    \mathbf{d^*_\BtoA} 
  \end{bmatrix}
\end{eqnarray}

Equation \ref{hppeval} clearly factors into the following two eigenvalue equations:
\begin{eqnarray}\label{b2aa2b}
    \kbordermatrix{
          & \BtoA              & \AtoB            \\
    \BtoA & \e\diag - \App     & -\Br             \\
    \AtoB & -\Br^*           & \e\diag - \App^* \\
  }
  \begin{bmatrix}
    \mathbf{d_\BtoA} \\
    \mathbf{d^*_\AtoB} \\        
  \end{bmatrix} 
  &=& 2E_2
  \begin{bmatrix}
    \mathbf{d_\BtoA} \\
    \mathbf{d^*_\AtoB} \\        
  \end{bmatrix}
  \\
  \label{a2bb2a}
    \kbordermatrix{
          & \AtoB              & \BtoA            \\
    \AtoB & \e\diag - \App     & -\Br             \\
    \BtoA & -\Br^*           & \e\diag - \App^* \\
  }
  \begin{bmatrix}
    \mathbf{d_\AtoB} \\
    \mathbf{d^*_\BtoA} \\        
  \end{bmatrix} 
  &=& 2E_2
  \begin{bmatrix}
    \mathbf{d_\AtoB} \\
    \mathbf{d^*_\BtoA} \\        
  \end{bmatrix}
\end{eqnarray}

Equations \ref{b2aa2b} and \ref{a2bb2a} are complex conjugates in the general case. If however, 
the solution is an RHF solution, the spatial orbitals for $\alpha$ and $\beta$ spins are identical
by definition. In this case, the matrices in both equation \ref{b2aa2b} and equation \ref{a2bb2a}
are equivalent to the triplet instability matrix defined by Seeger and Pople:

\begin{eqnarray}
  \mathbf{{}^3H} =
  \begin{bmatrix}
    \e \diag - (aj|bi) & -(ab | ji )           \\
    -(ab | ji )^*        & \e \diag - (aj|bi)^* \\        
  \end{bmatrix} 
\end{eqnarray}

The implication of equations \ref{b2aa2b}, \ref{a2bb2a} and \ref{3Heval} is that the following 
sets of eigenvectors comprise a degenerate manifold when the solution is an RHF wavefunction.
\begin{eqnarray}
  \begin{bmatrix}
    \mathbf{d_\BtoA} \\
    \mathbf{d^*_\AtoB} \\        
  \end{bmatrix}
  , 
  \begin{bmatrix}
    \mathbf{d_\AtoB} \\
    \mathbf{d^*_\BtoA} \\        
  \end{bmatrix}
  ,
  \begin{bmatrix}
    \mathbf{d_\AtoA - d_\BtoB} \\
    \mathbf{d^*_\AtoA - d^*_\BtoB} \\        
  \end{bmatrix}
\end{eqnarray}
As such, the RHF-GHF instability eigenvectors can always be transformed into eigenvectors having the
form of the RHF-UHF instability. On the other hand, if the solution is not an RHF solution, this is
not the case and the eigenvalues/vectors and corresponding instabilities are distinct.

\section{Numerical result}
To verify numerically, the diagonalization of the matrices $\mathbf{H}, \mathbf{H}'$ and 
${}^3\mathbf{H}'$ were implemented independently of one another. No factorizations or 
simplifications were performed. The lowest eigenvalue of each is shown below for a constant number
of K-points and a varying density. The results are equivalent within the convergence criteria of 
Davidson's Algorithm.  
\begin{figure}[h]
  \includegraphics[width=0.75\linewidth]{../../images/Matrix_Compare.pdf}
\end{figure}

\begin{comment}
\section{Proof of Real-Valued A and B}
In the case of the Homogeneous electron gas,
the two electron integral is given by
(eq. 12 of~\cite{Delyon2008} and p. 16 of~\cite{Guiliani2005}):

@@

So far I have not used this to simplify anything, but it is true.
\end{comment}
\section{References}
\bibliography{../heg_references}
\end{document}
